---
title: "HW12_ggplot"
author: "Owen R. Page"
date: "April 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will be using the data set ```iris```

Preliminaries:
```{r}
library(ggplot2)
library(patchwork)
library(ggthemes)
library(TeachingDemos)
char2seed("petrel and the porpoise")
```

The first graph is the smoothed regression lines by species for the continuous variables of sepal width and petal length:
```{r}
g1 <- ggplot(data=iris, mapping=aes(x=Sepal.Width, y=Petal.Length, col=Species))+
  geom_smooth(method="lm", se=FALSE)+
  theme_economist_white(base_family="serif")+
  geom_point()+
  xlab("Sepal Width")+
  ylab('Petal Length')


g1
```

The second graph is more simply a boxplot by species of petal length with the data points present:
```{r}
g2 <- ggplot(data=iris, mapping=aes(x=Species, y=Petal.Length, fill=Species))+
  theme_classic(base_family="serif")+
  geom_boxplot()+
  geom_point(position=position_jitter(width=0.2, height=0.7), color="grey30", alpha=0.4)
g2
```

The third graph is a histogram of petal width further broken up for each species: 
```{r}
g3 <- ggplot(data=iris, mapping=aes(x=Petal.Width, fill=Species))+
  geom_histogram(position='Dodge')+
  theme_economist(base_family="serif")
g3
```

Finally the fourth graph is of two continuous variables, sepal width and sepal length with a third variable included via a colour gradient:
```{r}
g4 <- ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width, color=Petal.Length))+
  geom_point()+
  scale_fill_brewer(palette="Blues")+
  theme_dark(base_family="serif")
g4
```


I used patchwork to put all four graphs together in two rows. I elected to make the second column wider as I believed this suited the data and make it easier to visualize:
```{r}
g2 + g3 + g1 +g4 + plot_layout(ncol=2, widths=c(1,2))
```

