---
title: "HW_10_BatchProcessing"
author: "Owen R. Page"
date: "March 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here is my code for Batch Processing

```{r, warning=FALSE}
##################################################
# function: FileBuilder
# create a set of random files for regression
# input: fileN = number of files to create
#       : fileFolder = name of folder for random files
#       : fileSize = c(min,max) number of rows in file
#       : fileNA = number on average of NA values per column
# output: set of random files
#------------------------------------------------- 
FileBuilder <- function(fileN=10,
                        fileFolder="RandomFiles/",
                        fileSize=c(15,100),
                        fileNA=3){
  for (i in seq_len(fileN)) {
    fileLength <- sample(fileSize[1]:fileSize[2],size=1) # get number of rows
    varX <- runif(fileLength) # create random x
    varY <- runif(fileLength) # create randon y
    dF <- data.frame(varX,varY) # bind into a data frame
    badVals <- rpois(n=1,lambda=fileNA) # determine NA number
    dF[sample(nrow(dF),size=badVals),1] <- NA # random NA in varX
    dF[sample(nrow(dF),size=badVals),2] <- NA # random NA in varY
    
    # create label for file name with padded zeroes
    fileLabel <- paste(fileFolder,
                       "ranFile",
                       formatC(i,
                               width=3,
                               format="d",
                               flag="0"),
                       ".csv",sep="")
    
    # set up data file and incorporate time stamp and minimal metadata
    write.table(cat("# Simulated random data file for batch processing","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# ORP","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=fileLabel,
                    row.names="",
                    col.names="",
                    sep=""))
    
    # now add the data frame
    write.table(x=dF,
                file=fileLabel,
                sep=",",
                row.names=FALSE,
                append=TRUE)
    
    
  }
}

##################################################
# function: regStats
# fits linear model, extracts statistics
# input: 2-column data frame (x and y)
# output: slope, p-value, and r2
#------------------------------------------------- 
regStats <- function(d=NULL) {
  if(is.null(d)) {
    xVar <- runif(10)
    yVar <- runif(10)
    d <- data.frame(xVar,yVar)
  }
  . <- lm(data=d,d[,2]~d[,1])
  . <- summary(.)
  statsList <- list(Slope=.$coefficients[2,1],
                    pVal=.$coefficients[2,4],
                    r2=.$r.squared)
  return(statsList)
  
}

##----------- Body for actual Batch Processing
library(TeachingDemos)
char2seed("Freezing March")

#--------------------------------------------
# Global variables
fileFolder <- "RandomFiles/"
nFiles <- 100
fileOut <- "StatsSummary.csv"
#--------------------------------------------

# Create 100 random data sets
FileBuilder(fileN=nFiles)

fileNames <- list.files(path=fileFolder)

# Create data frame to hold file summary statistics
ID <- seq_along(fileNames)
fileName <- fileNames
slope <- rep(NA,nFiles)
pVal <- rep(NA,nFiles)
r2 <- rep(NA,nFiles)

statsOut <- data.frame(ID,fileName,slope,pVal,r2)

# batch process by looping through individual files
for (i in seq_along(fileNames)) {
  data <- read.table(file=paste(fileFolder,fileNames[i],sep=""),
                     sep=",",
                     header=TRUE) # read in next data file
  
  dClean <- data[complete.cases(data),] # get clean cases
  
  . <- regStats(dClean) # pull regression stats from clean file
  statsOut[i,3:5] <- unlist(.) # unlist, copy into last 3 columns
  
}
# set up output file and incorporate time stamp and minimal metadata
write.table(cat("# Summary stats for ",
                "batch processing of regression models","\n",
                "# timestamp: ",as.character(Sys.time()),"\n",
                "# ORP","\n",
                "# ------------------------", "\n",
                "\n",
                file=fileOut,
                row.names="",
                col.names="",
                sep=""))

# now add the data frame
write.table(x=statsOut,
            file=fileOut,
            row.names=FALSE,
            col.names=TRUE,
            sep=",",
            append=TRUE)
```


4)  One possible manner by which to break the code is make the value for fileNA larger than the amount of rows (in this case, fileSize=c(15,100) so I altered fileNA to 16. This ended up breaking the program resulting in the following error:

Error in sample.int(x, size, replace, prob) : 
  cannot take a sample larger than the population when 'replace = FALSE'
 
  This program does still create random files in the RandomFile data folder--it only will create fifteen files--yet it does not create a Stats Summary .csv. 
  
 Another possible error that is more directly involved in the linear fit model:
we first altered the fileNA to 10 with fileSize=c(15,100). In the for loop below this we altered the rpois function and found that if rpois(n=1, lambda=fileNA) has n= altered to an odd number 15 or higher that the following error occurs:

Error in lm.fit(x, y, offset = offset, singular.ok = singular.ok, ...) : 
  0 (non-NA) cases

Here we find that the program creates all of the data in the RandomFiles yet it does not form the summary statistics .csv.

5) Altered the code so my summary statistics also has the number of original rows and rows in the cleaned data:

```{r}
for (i in seq_along(fileNames)){
  data <- read.table(file = paste(fileFolder, fileNames[i], sep=""),
                     sep=",",
                     header = TRUE) #read in next data file
  dClean <- data[complete.cases(data),] #get clean cases 
  
  . <- regStats(dClean) # pull regression stats from clean file
  statsOut[i, 3:5] <- unlist(.) #unlist, copy into last 3 columns
  statsOut[i,6] <- nrow(data) # new line
  statsOut[i,7] <- nrow(dClean)# new line
}

write.table(cat("# Summary stats for ", 
                "batch processing of regression models", "\n",
                "# timestamp: ",as.character(Sys.time()), "n", 
                "# ORP", "\n",
                "# --------------", "\n",
                "\n",
                file = fileOut, 
                row.names = "", 
                col.names = "", 
                sep = ""))
head(statsOut)
```
From this we can see that the cleaned row--V7--has lower values.

